Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-09-15 paper/agree-semantics.tex
Revision to make paper flow with overview

\begin{comment}
This section gives an abstract mathematical overview of AGREE's
compositional reasoning system, defining assume-guarantee contracts
and the verification conditions that AGREE checks in order to
modularly establish overall system correctness at the model level.
The syntax and semantics of AGREE are sketched in some detail.  The
discussion then becomes more concrete, showing---by example---how the
AGREE contract language is used to specify the cyber-hardened system
of \figref{fig:hardened}. The discussion proceeds from the
system-level contract to the filter and monitor components, which rely
on the notion of a \emph{code contract} to support code generation.
\end{comment}

We now give an overview of a formal model for AGREE syntax and
semantics.  The discussion then becomes more concrete, showing by
example, how the AGREE language is used to specify the properties of
the SW system introduced in \secref{sec:example}.  Any implementation
of the SW system must guarantee the specification output under the
system assumptions.

The syntax of AGREE is essentially that of quantifier-free first order
predicate logic supplemented with a few temporal operators. The terms
(\emph{e}) are arithmetic expressions built from variables ($v$) and
numeric and boolean literals ($c$), while formulas (\emph{b}) are
built using logical connectives from atomic formulas (\emph{a}) based
on the familiar comparison operators.
\[
\begin{array}{rcl}
e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
a & ::= & e\; \set{=,<}\; e \\
b & ::= & v \mid c \mid a \mid \neg b
            \mid b \; \set{\land,\lor,\imp,\iff}\; b
\end{array}
\]
There is also a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
and formulas. To support port handling, there is a predicate
$\konst{event}(p)$, which tests if an event (or event-data) port $p$ is
signaled. Lastly, there are temporal operators \emph{previously}
$\konst{pre}(-)$, \emph{followed-by} $(-) \to (-)$, and historically
$\konst{Hist}(-)$.

\subsection{Semantics}
The semantics of terms and formulas, as mentioned in the previous
section, is in terms of \emph{streams of values}. Values encompass at
least booleans and numbers, but can be readily extended to include
records and arrays. A value stream is a total function from time
(natural numbers including 0) to values:
\[
 \konst{stream} = \mathbb{N}_0 \to \konst{value}
\]
Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
binding variable names to value streams, the semantics $\sem{-}^E_t$
of terms and formulas defines the meaning of compound syntax in terms
of the meaning of subexpressions. The value of a variable $v$ at time
$t$ is found by looking up the stream bound to $v$ in $E$ (call it
$s$) and returning $s_t$. Some clauses of the semantics follow,
omitting the temporal operators:
\[
\begin{array}{rcl}
\sem{v}^E_t & = & E(v)(t) \\
\sem{c}^E_t & = & c \\
\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
   & \cdots & \\
\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
   & \cdots & \\
\end{array}
\]

@@

\paragraph{AGREE notation.}
AGREE uses ASCII notation for the mathematical syntax we have
defined. The following table spells out the mapping.
\[
\begin{tabular}{l|l|l}
 Operation & Math. notation & AGREE \\ \hline
 followed-by &  $e \to e$          & e \verb+->+ e  \\
 logical implication &  $e \imp e$ & e \verb+=>+ e \\
 if-and-only-if &  $e \iff e$      & e \verb+<=>+ e \\
 variable definition & $v = e$     & \konst{eq} v\ e \\
 Historically  & $\konst{Hist}(b)$ & \konst{Historically}(b) \\
\end{tabular}
\]

>>>>

The syntax of \agr\ is essentially that of quantifier-free first order
predicate logic supplemented with a few temporal operators. The terms
(\emph{e}) are arithmetic expressions built from variables ($v$) and
numeric and boolean literals ($c$), while formulas (\emph{b}) are
built using logical connectives from atomic formulas (\emph{a}) based
on the familiar comparison operators.
\[
\begin{array}{rcl}
e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
a & ::= & e\; \set{=,<}\; e \\
b & ::= & v \mid c \mid a \mid \neg b
            \mid b \; \set{\land,\lor,\imp,\iff}\; b
\end{array}
\]
\agr\ includes a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
and formulas. To support port handling, \agr\ defines a predicate
$\konst{event}(p)$, which tests if an event (or event-data) port $p$ is
signaled. Lastly, \agr\ supports temporal operators \emph{previously}
$\konst{pre}(-)$, \emph{followed-by} $(-) \to (-)$, and historically
$\konst{Hist}(-)$.

\subsection{Semantics}
The semantics of terms and formulas is defined over \emph{streams of values}. Values encompass at
least booleans and numbers but can be readily extended to include
records and arrays. A value stream is a total function from time
(natural numbers including 0) to values:
\[
 \konst{stream} = \mathbb{N}_0 \to \konst{value}
\]
Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
binding variable names to value streams, the semantics $\sem{-}^E_t$
of terms and formulas defines the meaning of compound syntax in terms
of the meaning of subexpressions. The value of a variable $v$ at time
$t$ is found by looking up the stream bound to $v$ in $E$ (call it
$s$) and returning $s_t$. Some clauses of the semantics follow,
omitting the temporal operators:
\[
\begin{array}{rcl}
\sem{v}^E_t & = & E(v)(t) \\
\sem{c}^E_t & = & c \\
\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
   & \cdots & \\
\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
   & \cdots & \\
\end{array}
\]

@@

\agr\ uses ASCII notation for the mathematical syntax we have
defined. The following table spells out the mapping.
\[
\begin{tabular}{l|l|l}
 Operation & Math. notation & \agr \\ \hline
 followed-by &  $e \to e$          & \verb+e -> e+  \\
 logical implication &  $e \imp e$ & \verb+e => e+ \\
 if-and-only-if &  $e \iff e$      & \verb+e <=> e+ \\
 logical and &  $e \land e$      & \verb+e and e+ \\
 logical or &  $e \lor e$      & \verb+e or e+ \\
 logical not & $\neg b$      & \verb+not b+ \\
 variable definition & $v = e$     & \verb+eq v e+ \\
 Previously  & $\konst{pre}(b)$ & \verb+pre(b)+ \\
 Historically  & $\konst{Hist}(b)$ & \verb+Historically(b)+ \\
\end{tabular}
\]

### 2022-09-15 paper/example.tex
Revision to make paper flow with overview

The baseline AADL model is a simple software system, named SW, for route planning
and automated control of a UAV. A picture of the AADL architectural
model for SW is in \figref{fig:example}. SW is loosely based on one of
the case studies in Section~\ref{sec:case-study}.  The source for the
entire model is found at \cite{repo}.

From time to time, SW receives a message on the \texttt{AutomationRequest} input,
forwarding it to a route planner (AI). The function of AI is to 
compute the flight path (a list of waypoints) for the UAV, outputting
the resulting \emph{mission command} on \texttt{AutomationResponse}.
The waypoint manager (WM) receives the mission command from AI and
starts the UAV flying the mission by putting an \emph{event} on the
\texttt{Start} output port, continuing to issue waypoints to the UAV
flight controller on the \texttt{Waypoint} port as the UAV location
updates with messages on \texttt{AirVehicleLocation}.
An event is repeatedly generated on \texttt{Alert} if there is no
mission command from a request. The AI component is third-party
software and WM is a legacy component.

@@

The expected behavior of the SW system, and the components
implementing the system, are modeled with \agr\ contract
specifications.  
These contracts assume and guarantee the absence
of any malicious, or unspecified, component behavior.  More precisely,
the contract for SW assumes that its inputs are \emph{well-formed} and that
there is never more than one automation request pending at a time.
Well-formed generally refers to a syntactic restriction on
data at a port. For example, a waypoint is well-formed if it falls
within bounds for latitude, longitude, and altitude.  The guarantees
for SW ensure that
\begin{compactitem}
\item a start coincides with a new waypoint message being output;
\item a start is within one cycle of an automation request and if not, then it persistently alerts;
\item new waypoints coincide with air vehicle location updates; and
\item all outputs are well-formed.
\end{compactitem}

The contracts for the sub-components assume their inputs are
well-formed, and they guarantee their outputs are well-formed.  The AI
contract guarantees it only responds to automation requests and always
in the same cycle.  The legacy WM contract guarantees the following if
its input assumptions hold:
\begin{compactitem}
  \item it generates a start from a response from the AI and always in the same cycle;
  \item the start always coincides with a waypoint being output; and
  \item any further waypoints coincide with air vehicle location updates.
\end{compactitem}
These initial specifications pass verification, meaning that the
contract composition of the components with the system satisfies all
component input assumptions and system output guarantees.
The \agr\ verification conditions, syntax and semantics, and the baseline contracts for the example are discussed in detail in \secref{sec:agree} and \secref{sec:agree-semantics}.

@@

A cyber-vulnerability analysis identifies the potential of a
supply chain attack through the AI route planner since its third-party.  Based on the analysis, the designer adds a requirement that the system must guard against malicious AI behavior, marks the component as \emph{untrusted}, and removes all output guarantees from its contract.  The behavior of AI is now unconstrained.

@@

Here is where the shaded regions in \figref{fig:flowchart} come into play.
The system designer now uses \brfcs\ to cyber-harden SW by inserting
high-assurance components in the form of a filter and a monitor, as
shown in \figref{fig:hardened}.
These are intended to mitigate and report any observed malicious behavior by AI.
A filter enforces an invariant over
each datum in the data stream by not forwarding input to its output if the invariant does not hold.
\brfcs\ generates an initial code contract for the filter which the designer completes by writing the invariant.
That invariant is usually based on the existing assumptions made by
downstream components that consume the filter output.
In this example, it is the well-formed property for the automation request.

@@

Code contracts for high assurance components can be arbitrarily complex since they can have persistent state that evolves over time.
Test contracts allow the designer to validate their behavior with unit testing.
For example, one of the test contracts for the monitor is that when it sees an automation response one step after a request, then it should not alert, and it should pass the response downstream.
Test contracts are discussed in detail in \secref{sec:test-contracts}.

@@

High-assurance components are automatically synthesized by \splt\ from
the code contracts to equivalent programs in the \ckml\
language.
In other words, for any set of input streams that meet the
component's contract assumptions, the output streams produced from
the synthesized \ckml\ code exactly match those defined by the contract guarantees.
As such, the \agr\ verification results regarding the code contract apply equally to the synthesized \ckml, and thus, the resulting binaries from the \ckml\ compiler.
This result does not extend to the whole system composition.
Synthesis from code contracts is discussed in detail in \secref{sec:sythesis}.

Extending the \agr\ verification results to the system composition requires additional guarantees not discussed in this manuscript.
For example, among other things, contracts for non-synthesized components must be certified to accurately model their deployed counterparts, and the HAMR generated scheduler follows the dependent data-flow in the design.
These additional guarantees are discussed in other works \cite{gearcase2020, dcrypps2019, 10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17, sel4-2009, scheduled-agree, 9734792}.

>>>>

The baseline AADL model is a simple software system, named SW, for route planning
and automated control of a UAV. A picture of the architectural
model for SW is in \figref{fig:example}. SW is loosely based on one of
the case studies in Section~\ref{sec:case-study}.  The source for the
entire model is found at \cite{repo}.

From time to time, SW receives a message on the \texttt{AutomationRequest} input,
forwarding it to a route planner (AI). The function of AI is to 
compute the flight path (a list of waypoints) for the UAV, outputting
the resulting \emph{mission command} on \texttt{AutomationResponse}.
The waypoint manager (WM) receives the mission command from AI and
starts the UAV flying the mission by putting an \emph{event} on the
\texttt{Start} output port, continuing to issue waypoints to the UAV
flight controller on the \texttt{Waypoint} port as the UAV location
updates with messages on \texttt{AirVehicleLocation}.
%%An event is repeatedly generated on \texttt{Alert} if there is no
%%mission command from a request. 
The AI component is third-party
software and WM is a legacy component.

@@

The expected behavior of the SW system, and the components
implementing the system, are modeled with \agr\ contract
specifications.  
These contracts assume and guarantee the absence
of any malicious, or unspecified, component behavior.  More precisely,
the contract for SW assumes that its inputs are \emph{well-formed} and that
there is never more than one automation request pending at a time.
Well-formed generally refers to a syntactic restriction on
data at a port. For example, a waypoint is well-formed if it falls
within bounds for latitude, longitude, and altitude.  The guarantees
for SW ensure that
\begin{compactitem}
\item a start coincides with a new waypoint being output;
\item a start is within one cycle of an automation request and if not, then it persistently alerts;
\item new waypoints coincide with location updates; and
\item all outputs are well-formed.
\end{compactitem}

The contracts for the sub-components assume their inputs are
well-formed, and they guarantee their outputs are well-formed.  The AI
contract guarantees it only responds to automation requests and always
in the same cycle.  The legacy WM contract guarantees that
\begin{compactitem}
  \item it generates a start from a response from the AI and always in the same cycle;
  \item the start coincides with a waypoint; and
  \item any further waypoints coincide with location updates.
\end{compactitem}

These initial specifications pass verification, meaning that the
contract composition of the components with the system satisfies all
component input assumptions and system output guarantees.
The \agr\ verification conditions, syntax and semantics, and the baseline contracts for the example are discussed in detail in \secref{sec:agree}.

@@

A cyber-vulnerability analysis identifies the potential of a
supply chain attack through the AI route planner since its third-party.  Based on the analysis, the designer adds a requirement that the system must guard against malicious AI behavior.
They additionally remove all output guarantees from its contract since it is untrusted.

@@

Here is where the shaded regions in \figref{fig:flowchart} come into play.
The system designer now uses \brfcs\ to cyber-harden SW by inserting
high-assurance components in the form of a filter and a monitor, as
shown in \figref{fig:hardened}.
These are intended to mitigate and report any observed malicious behavior by AI.

A filter enforces an invariant over
each datum in the data stream by not forwarding input to its output if the invariant does not hold.
\brfcs\ generates an initial code contract for the filter which the designer completes by writing the invariant.
That invariant is usually based on the existing assumptions made by
downstream components that consume the filter output.
In this example, it is the well-formed property for the automation request.

@@

Code contracts for high assurance components can be arbitrarily complex since they can have persistent state that evolves over time.
Test contracts allow the designer to validate code contract behavior with unit testing.
For example, one of the test contracts for the monitor is that when it sees an automation response one step after a request, then it should not alert, and it should pass the response downstream.
Test contracts are discussed in detail in \secref{sec:testing}.

@@

High-assurance components are automatically synthesized by \splt\ from
the code contracts to equivalent programs in the \ckml\
language.
In other words, for any set of input streams that meet the
component's contract assumptions, the output streams produced from
the synthesized \ckml\ code exactly match those defined by the contract guarantees.
As such, the \agr\ verification results regarding the code contract apply equally to the synthesized \ckml, and thus, the results equally apply to the resulting binaries from the \ckml\ compiler.
Synthesis from code contracts is discussed in detail in \secref{sec:synthesis}.

Extending the \agr\ verification results to the system composition requires additional guarantees not discussed in this manuscript.
For example, among other things, contracts for non-synthesized components must be certified to accurately model their deployed counterparts, the HAMR generated scheduler must be certified to follow the dependent data-flow in the design, scheduling windows must be certified to cover worst-case execution times, etc.
Such additional guarantees are discussed in other works \cite{gearcase2020, dcrypps2019, 10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17, sel4-2009, scheduled-agree, 9734792}.

### 2022-09-15 paper/high-assurance-contracts.tex
Revision to make paper flow with overview

The semantics of Section \ref{agree-semantics} supports a formal
connection between the original contract---and verification results of
the AGREE model checker being run on it---and code generated from
specifications of leaf level cyber-components. It also provides the
root meaning at the base of a chain of translation steps moving from
an AGREE contract to a CakeML executable.

>>>>

The semantics of \secref{sec:agree-semantics} supports a formal
connection between the original contract---and verification results of
the AGREE model checker being run on it---and code generated from
specifications of leaf level cyber-components. It also provides the
root meaning at the base of a chain of translation steps moving from
an AGREE contract to a CakeML executable.

### 2022-09-15 paper/overview.tex
Revision to make paper flow with overview

\brfcs\ integrates additional tools into OSATE to analyze, identify, and address cybersecurity vulnerabilities in the baseline AADL model (top of \figref{fig:tool-arch}).
\brfcs\ provides access to analysis tools that examine the model, report potential cyber vulnerabilities,
and suggest requirements for mitigation \cite{dcrypps2019,gearcase2020}.
Designers add these requirements to the \brfcs\ project requirements repository. 
A key part of the requirements management in \brfcs\ is a corresponding assurance case that provides evidence, computed directly from the model or by supporting analysis tools, showing how each requirement is satisfied.
Evidential statements must exist to address the added cyber requirements in order for \brfcs\ to build the assurance case for the project.

\brfcs\ provides an extendable library of model transformations to address requirements for common cyber vulnerabilities:
\begin{compactitem}
	\item \textbf{Filter}: block messages that do not conform to a given specification
	\item \textbf{Monitor}: detect and alert violations at runtime
	\item \textbf{Gate}: block messages in the presence of an alert
	\item \textbf{Attestation}: measure non-local software to assess trustworthiness
	\item \textbf{Virtualization}: isolate components with virtualization
	\item \textbf{Proxy}: enable HTTPS message inspection 
	\item \textbf{seL4}: transform component definitions for the seL4 secure micro kernel
\end{compactitem}
Transformations update the AADL model architecture with new components and the assurance case with new required evidential statements.

@@

The Awas AADL information flow analyzer and visualizer provides evidence that high assurance components are connected and not bypassed in the model \cite{awas}.
The High Assurance Modeling and Rapid engineering for embedded systems tool (HAMR) generates from the model the component infrastructure and integration code that implements the semantics of AADL-compliant thread scheduling, thread dispatching, port-based communication, and real-time scheduling \cite{hamr}.
Along with that code generation, HAMR constructs a proof that all flows in the model are in the implementation and that no extraneous flows have been added to the implementation.
HAMR relies on the seL4 microkernel for its proofs.
The seL4 microkernel, among other security properties, provides strong isolation between components ensuring that there are no communication channels outside of what is defined \cite{sel4-sosp09, sel4-tocs14, sel4-cacm18}.

The Assume Guarantee Reasoning Environment (\agr) is a compositional, assume-guarantee-style, model checker for AADL models \cite{compositional-analysis-agree,nfm:agree}. 
\agr\ attempts to prove properties about one layer of the model using properties allocated to sub-components.
The composition is performed in terms of \emph{contracts} provided for each component and sub-component that state \emph{assumptions} on inputs and \emph{guarantees} on outputs under the assumptions.
\agr\ verifies the \emph{consistency} of the composition of the component and sub-component contracts meaning that in every connection the source constraints satisfy the destination constraints.

The contributions discussed in this manuscript relate to the aspects of \brfcs\ concerning evidence that a high-assurance component meets its intended purpose and is properly implemented.
This manuscript defines a \emph{code contract} language that is a subset of the \agr\ language with sequential evaluation semantics as opposed to the existing parallel semantics of \agr.
The sequential semantics are requisite to code synthesis and are defined such that \agr\ is able to still prove consistency in the contract composition of the model. 
This manuscript further defines a stylized \emph{test contract} for unit testing code contracts outside of the system composition to verify input to output behavior as well as general properties of the code contract.
As such, the behavior of a high assurance component is defined by the designer in a code contract, and then using the existing \agr\ model checker, test contracts prove the desired code contract behavior.
Once the designer is satisfied that the code contract is correct, the system composition is verified to conclude that the contract meets its intended purpose.

Showing that a component is properly implemented is accomplished through formal synthesis from the code contract (bottom of \figref{fig:tool-arch}).
The Semantic Properties of Language and Automata Theory (\splt) tool generates \emph{\ckml} code to implement the code contract.
\ckml\ is a functional language that includes additional proofs and tools built around it \cite{cakeml}.
The \ckml\ compiler itself is verified to generate binary implementations that are equivalent in semantics to the original input program.
That proof ensures that the resulting binary exactly preserves the meaning of the original code contract and assures that the component is properly implemented.

>>>>

\brfcs\ integrates additional tools into OSATE to analyze, identify, and address cybersecurity vulnerabilities in the baseline model (top of \figref{fig:tool-arch}).
\brfcs\ provides access to analysis tools that examine the model, report potential cyber vulnerabilities,
and suggest requirements for mitigation \cite{dcrypps2019,gearcase2020}.
Designers add these requirements to the \brfcs\ project requirements repository. 
A key part of the requirements management in \brfcs\ is a corresponding assurance case that provides evidence, computed directly from the model or by supporting analysis tools, showing how each requirement is satisfied.
Evidential statements must exist to address the added cyber requirements in order for \brfcs\ to build the assurance case for the project.

\brfcs\ provides an extendable library of model transformations to address requirements for common cyber vulnerabilities:
\begin{compactitem}
	\item \textbf{Filter}: block messages that do not conform to a given specification
	\item \textbf{Monitor}: detect and alert violations at runtime
	\item \textbf{Gate}: block messages in the presence of an alert
	\item \textbf{Attestation}: measure non-local software to assess trustworthiness
	\item \textbf{Virtualization}: isolate components
	\item \textbf{Proxy}: enable HTTPS message inspection 
	\item \textbf{seL4}: transform component definitions for the seL4 secure micro kernel
\end{compactitem}
Transformations update the model's architecture with new components and the assurance case with new required evidential statements.

@@

The Awas AADL information flow analyzer and visualizer provides evidence that high assurance components are connected and not bypassed in the model \cite{awas}.
The High Assurance Modeling and Rapid engineering for embedded systems tool (HAMR) generates from the model the infrastructure and integration code that implements the semantics of AADL-compliant scheduling, dispatching, and port-based communication \cite{hamr}.
HAMR additionally constructs a proof that all flows in the model are in the implementation and that no extraneous flows have been added.
HAMR relies on properties of the seL4 microkernel for some aspects of these proofs.
The seL4 microkernel, among other things, provides strong isolation between components ensuring that there are no communication channels outside of what is defined \cite{sel4-sosp09, sel4-tocs14, sel4-cacm18}.

Important to the contributions presented in this paper is the Assume Guarantee Reasoning Environment (\agr) \cite{compositional-analysis-agree,nfm:agree}. It is a compositional, assume-guarantee-style, model checker for AADL models. 
\agr\ attempts to prove properties about one layer of the model using properties allocated to sub-components.
The composition is performed in terms of \emph{contracts} provided for each component and sub-component that state \emph{assumptions} on inputs and \emph{guarantees} on outputs under the assumptions.
\agr\ verifies the \emph{consistency} of the composition of the component and sub-component contracts meaning that in every connection the source constraints satisfy the destination constraints.

The contributions discussed in this manuscript relate to the aspects of \brfcs\ concerning evidence that a high-assurance component meets its intended purpose and is properly implemented.
This manuscript defines a \emph{code contract} language that is a subset of the \agr\ language with sequential evaluation semantics as opposed to the existing parallel semantics of \agr.
The sequential semantics are requisite to code synthesis and are defined such that \agr\ is able to still prove consistency in the contract composition of the model.

This manuscript further defines a stylized \emph{test contract} for unit testing code contracts outside of the system composition to verify input to output behavior as well as general properties of the code contract.
As such, the behavior of a high assurance component is defined by the designer in a code contract, and then using the existing \agr\ model checker, test contracts prove the desired code contract behavior.
Once the designer is satisfied that the code contract is correct, the system composition is verified to conclude that the contract meets its intended purpose.

Showing that a component is properly implemented is accomplished through formal synthesis from the code contract (bottom of \figref{fig:tool-arch}).
This manuscript presents the Semantic Properties of Language and Automata Theory (\splt) tool that generates \emph{\ckml} code to implement the code contract.
\ckml\ is a functional language that includes additional proofs and tools built around it \cite{cakeml}.
The \ckml\ compiler itself is verified to generate binary implementations that are equivalent in semantics to the original input program.
That proof ensures that the resulting binary exactly preserves the meaning of the original code contract and assures that the component is properly implemented.

### 2022-09-15 paper/testing.tex
Revision to make paper flow with overview

Before discussing the final steps in synthesis, we take a moment to
discuss how code contracts, and any AGREE specification in general, can be traditionally tested to build assurance.
Writing and proving properties about AGREE specifications is hard not just because it is a formal process but  because it reasons about constraints that define the entire input space for a component, and the entire output space, at once.
The case studies showed that engineers, and formal method practitioners, make mistakes writing these specifications especially if the computation or temporal reasoning is complex.

We have observed three common types of mistakes in specifications (aside from self-consistency, which is
already checked): vacuity, under-specification, and missed corner cases.
Vacuity is a common pitfall with implication when the left hand side is always false, making the implication true. 
Under-specification is very challenging, as a verifier in the face of under-specification is not able to prove anything (or it can prove everything in the case of existential properties), so nothing is actually known after verification except that the system is under-specified. 
Missed corner cases is a common problem to all disciplines.
It is an especially difficult challenge here because there is no direct way to provide test input and check test output.

The tempting approach is to use the synthesis path to check these issues with traditional testing.
Such an approach adds extra steps that are not needed, as it is
possible to bridge the formal model in AGREE, with its proof system, to
traditional validation using tests.
The key is to consider the component as implementing a test specification.

Unit tests are created in AGREE by writing a system specification that assumes the test input and then guarantees the test output.
The system is then implemented with the component under test and its specification.
The AGREE analysis then proves if the component guarantees, given the test inputs, are strong enough to prove the guaranteed test outputs.
Here the system assumptions strengthen the component assumptions to a single input, and the system guarantees weaken, or leave unchanged, the component output for that test input, depending on the goal of the test.
AGREE proves that the component implements the test, and the Liskov substitution principle holds.

@@

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
    \scalebox{0.62}{\usebox{\tst}} \\
    \end{tabular}
  \end{center}
  \caption{AGREE unit test for high-assurance monitor.}
  \label{fig:test}
\end{figure}

\figref{fig:test} is a unit test for the high-assurance monitor in \figref{fig:hardened}.
The test checks if the monitor output coincides with the response that comes one step after the request.
It also checks that the alarm does not fire.
\begin{compactitem}
\item \lineref{line:tst-index}: An index to use to define the test input and expected test output through time.
\item \lineref{line:tst-assume-1} and \lineref{line:tst-assume-2}: The test input. These define when the request and response message events should appear relative to the index.
\item \lineref{line:tst-guarantee-1} and \lineref{line:tst-guarantee-2}: The expected test output. 
The test additionally proves that nothing happens after the initial test input and expected output, \ie, no more input is provided and no more output is generated.
\item \linesref{line:tst-imp-start}{line:tst-imp-end}: The implementation with the component under test (\lineref{line:tst-imp-comp}).
\end{compactitem}
\figref{fig:test-output} is the AGREE output for the test, showing that it passes.

This AGREE test method is effective for detecting vacuity, under-specified behavior, and missed corner cases.
It enables designers to iteratively develop specifications inside the MBSE framework using AGREE, and it builds assurance that the specification reflects the desired behavior with black-box, white-box, and other test coverage metrics.
These same tests can be synthesized to the backend implementation for traditional testing.

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.38]{agree-test-output.png} \\
    \end{tabular}
  \end{center}
  \caption{AGREE output for unit test.} 
  \label{fig:test-output}
\end{figure}

>>>>

Writing and proving properties about code contracts is difficult because it is a formal process, and because it reasons about constraints that define the entire input space, and the entire output space, of a high-assurance component at once.
Case studies have shown that designers, as well as formal method practitioners, make mistakes writing these contracts, especially if the computation or temporal reasoning in the contract is complex.

We have observed three common types of mistakes in code contracts (aside from self-consistency, which is
already checked): vacuity, under-specification, and missed corner cases.
Vacuity is a common pitfall with implication when the left hand side is always false, making the implication true. 
Under-specification is very challenging, as \agr\ is not able to prove anything (or it can prove everything in the case of existential properties), so nothing is actually known after verification except that the contract is under-specified. 
Missed corner cases is a common problem to all disciplines.
It is an especially difficult challenge here because there is no direct to exercise code contracts (e.g., there is no runtime \emph{per se}).

A tempting approach to exercising code contracts directly is to use \splt\ to create the target binary and then use that runtime with with traditional testing.
Such an approach adds extra steps, including test harneses, that are not needed, as it is
possible to bridge the formal model in \agr, with its proof system, to
traditional validation using tests.
The intuition for that bridge comes from the compositional reasoning in AADL and \agr. 
Effectively, a code contract for a high assurance component should be a valid implementation of a test scenario.
In other words, the test scenario, as expressed in a test contract, is the desired behavior of the system, and the code contract implements that behavior.

A test contract defines the test scenario for the code contract to implement.
It does this by assuming the test input and guarantying the test output akin to a unit test. 
The code contract implementation (e.g., the subject under test).
The \agr\ analysis then proves if the code contract, given the test inputs, is strong enough to prove the test contract outputs.
Here the test contract assumptions strengthen the code contract assumptions to a single input, and the test contract guarantees weaken, or leave unchanged, the code contract output for that test input, depending on the goal of the test.
In other words, according to Liskov substitution, \agr\ proves if the code contract is a safe substitution for the test contract.

@@

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
    \scalebox{0.60}{\usebox{\tst}} \\
    \end{tabular}
  \end{center}
  \caption{A unit test for the monitor with its test contract.}
  \label{fig:test}
\end{figure}

\figref{fig:test} is a unit test, with its test contract, for the monitor in \figref{fig:hardened}.
The test contract checks if the monitor output coincides with the response that comes one step after the request.
It also checks that the alarm is not output.
\begin{compactitem}
\item \lineref{line:tst-index}: An index to use to define the test input and expected test output through time.
\item \lineref{line:tst-assume-1} and \lineref{line:tst-assume-2}: The test input. These define when the request and response message events should appear relative to the index.
\item \lineref{line:tst-guarantee-1} and \lineref{line:tst-guarantee-2}: The expected test output. 
The test additionally proves that nothing happens after the initial test input and expected output, \ie, no more input is provided and no more output is generated.
\item \linesref{line:tst-imp-start}{line:tst-imp-end}: The implementation with the component under test (\lineref{line:tst-imp-comp}).
\end{compactitem}
\figref{fig:test-output} is the \agr\ output for the test, showing that it passes.

Test contracts are effective in detecting vacuity, under-specified behavior, and missed corner cases.
They enables designers to iteratively develop code contracts inside OSATE with \agr, and they build assurance that the code contract reflects the desired behavior with black-box, white-box, and other test coverage metrics.
These same tests can be synthesized to the backend implementation for traditional testing if for example, the code contract is a model of an existing legacy binary.
